---
title: "Vaccination rates"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(reshape2)
library(gt)

#load data 
measles <-read.csv("~/GitHub/TidyTuesday/2020-02-25/measles.csv")

#Subset data by state 
df_list<-split(measles, format(as.factor(measles$state)))

measles_df<-bind_rows(df_list,.id = "id")

measles$mmr<- abs(measles$mmr)
measles$overall<- abs(measles$overall)
measles$xrel<- abs(measles$xrel)
measles$xmed<- abs(measles$xmed)
measles$xper<- abs(measles$xper)

measles[is.na(measles)] = 0

vaccine_rate<- measles %>%
  group_by(state) %>%
  summarise(across(mmr:xper, mean, na.rm = TRUE))

measles_df$type<-as.factor(measles_df$type)
measles_df$mmr<- abs(measles_df$mmr)

knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
This data set is from the Wall Street Journal and includes measles vaccination rates in schools across 32 states in the US. The data covers different years: vaccination rates are for the 2017-18 school year for Colorado, Connecticut, Minnesota, Montana, New Jersey, New York, North Dakota, Pennsylvania, South Dakota, Utah and Washington. Rates for other states are 2018-19.I wanted to try the gt package to summarize this data in tables. 

The table below shows the vaccination rates for each state, including the percent rate of exemptions for personal, medical, and religious reasons. For some entries zeros appear because the original data either had negative numbers or was missing. The latter mainly affects the exemptions numbers.
<div style="margin-bottom:30px;">
</div>

```{r plot-vaccinetable, echo=FALSE}
vaccine_rate %>% 
  gt()%>%
  tab_header(title = md("**Vaccination rate per state**"))%>%
  fmt_number(
    columns = vars(mmr,overall,xrel,xmed,xper),
    decimals = 2
  )%>%
tab_source_note(
    source_note = "Data Source: Wall Street Journal"
  ) %>%
cols_label(
  state = "State",
  mmr = "MMR vaccination rate",
  overall = "Overall vaccination rate",
  xrel = "Religious exemption",
  xmed = "Medical exemption",
  xper = "Personal exemption"
)
```
<div style="margin-bottom:30px;">
</div>

The next step is to calculate the vaccination rates per school type for each state. I was curious as to whether there were differences depending on school type.

```{r plot-vaccinerates-perschooltype, echo=FALSE}
schoolvaccinerate <-measles_df %>%
  group_by(as.factor(id),type) %>%
  summarize(mmr = mean(mmr,na.rm = TRUE,groups='drop'))

schoolvaccinerate %>% 
  gt()%>%
  tab_header(title = md("**Vaccination rate per school type**"))%>%
  fmt_number(
    columns = vars(mmr),
    decimals = 1
  )%>%
tab_source_note(
    source_note = "Data Source: Wall Street Journal"
  ) %>%
cols_label(
  'as.factor(id)' = "State",
  type = "School type",
  mmr = "Measles vaccination rate",
) %>%
fmt_missing(schoolvaccinerate,
    columns = 1,
    missing_text = "--")
```
<div style="margin-bottom:30px;">
</div>

Unfortunately, the school type i.e. Public, Private, and Charter schools is not included for all states. So the table output from the analysis includes NA values. In addition, some states include additional designations for school type in this data set, e.g. Kindergarten, BOCES etc

<!--Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.--->
